---
title: "Occupancy Script Rodent Cambodia"
author: "Mathieu Pruvot"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r librairies}

library(AHMbook)
library(tidyverse)
library(reshape2)
library(unmarked)
library(jagsUI)
library(vegan)
library(ggpubr)
library(ggplot2)
library(ggthemes)
```

```{r load data}
load("captanim12.Rdata")
load("traplayout_tot2.Rdata")
head(captanim12)
```



# Data preparation: convert capture data to occupancy data

Create an empty dataframe based on traplayout_tot2 and usage to list all trap/occasions in long form
```{r}
occuframe<-traplayout_tot2
occuframe$X1<-NA
occuframe$X2<-NA
occuframe$X3<-NA
occuframe$X4<-NA
occuframe$X5<-NA
occuframe$X6<-NA
occuframe$X7<-NA
occuframe$X8<-NA

t(as.data.frame(strsplit(as.character(occuframe$usage),"")))->occuframe[,c("X1","X2","X3","X4","X5","X6","X7","X8")]

```

Convert to long format
```{r}
meltoccuframe<-melt(occuframe, id.vars=c("TrapID","X","Y","usage","Site","season"  
,"Zone","Trap","SiteSeason"),variable_name = "occasion")
meltoccuframe<-meltoccuframe[order(meltoccuframe$season,meltoccuframe$Site,meltoccuframe$Zone,meltoccuframe$TrapID),]

meltoccuframeused<-meltoccuframe[meltoccuframe$value==1,]

```

replicate resulting dataframe by number of species

```{r}
nspe<-length(unique(captanim12$species))
speoccup<-purrr::map_dfr(seq_len(nspe), ~meltoccuframeused)
ntrapocc<-nrow(meltoccuframeused)
spetemp<-rep(unique(captanim12$species),ntrapocc)
spetemp<-spetemp[order(spetemp)]
speoccup$species<-as.character(spetemp)

```

create detection/non-detection and assign 0 to all

```{r}
speoccup$detection<-0

```



gsub X out of the speoccup occasion

```{r}
speoccup$variable<-gsub("X",replacement = "",x=speoccup$variable)

```

## transfer captanim12 data into speoccup df to document detections


First create the varialbe
```{r}

captvarname <- colnames(captanim12)[c(2,6,11:40)]
speoccup[captvarname] <- NA
```

Then, convert Occasion in captanim12 into absolute values (remove the - for mortality events)
```{r}
captanim12$Occasion <- abs(captanim12$Occasion)
```


then at each i, add the entire row to the list of variables in the bracket (the session serves as 1 for detection)
```{r}
for(i in 1:nrow(captanim12)){
  speoccup[speoccup$season==captanim12[i,]$Season.x & speoccup$TrapID==captanim12[i,]$TrapID & speoccup$variable==as.character(captanim12[i,]$Occasion) & speoccup$species==as.character(captanim12[i,]$species),13:45]<-captanim12[i,c(1:2,6,11:40)]
}
```

check that all Ids from captanim were transfered and that it generally looks good
```{r}
nrow(captanim12)-nrow(speoccup[speoccup$detection==1,]
)
captanim12[!captanim12$Id %in% speoccup$Id,]



```
OK!

# Community occupancy

```{r}
spespread<-speoccup[,c(1,5:10,12:13)]
spespread_tot <- spread(spespread,variable,detection)
```



## format data for community analysis
```{r}
head(spespread_tot,n=20)
```


```{r}
spespread_tot$sitezone<-paste(spespread_tot$SiteSeason,spespread_tot$Zone,sep="")
```

prepare array of site x trap x occasion x species 
```{r}
nsites<-3*9*2
ntraps <- 50 # number of traps
nreps <- 8 # max number of replicate surveys per trap
nspec <- length(unique(speoccup$species)) # 14 species occur in the data
ms_sp_a <- array(NA, dim = c(nsites, ntraps,nreps, nspec))
```


```{r}
(ordered.spec.name.list <- unique(speoccup$species)) # ID-order list
sitezone.list<-unique(spespread_tot$sitezone)


for(k in 1:nspec){
  for(i in 1:nsites)
    ms_sp_a[i,,,k] <- as.matrix(spespread_tot[spespread_tot$species==ordered.spec.name.list[k] & spespread_tot$sitezone==sitezone.list[i],8:15],ntraps)
}
dimnames(ms_sp_a) <- list(trapid=sitezone.list,trap=paste("T",as.character(1:50),sep=""), occasion=as.character(1:8),sps=ordered.spec.name.list)

```
# before doing the occupancy model, simple diversity estimate
now that we have the ms_sp_a array. we can easily collapse the array to the desired dimension. Let's start with the simplest approach possible, which is to produce a site occupancy matrix (site x species) to calculate simple diversity estimates
```{r}
divmat <- apply(ms_sp_a, c(1,4), sum, na.rm = T)
divmat01 <- divmat
divmat01[divmat01>0] <- 1


diversity(divmat)
zone <- rep(c("Z1","Z2","Z3"),18)
seas <- rep(c("d","r"),each=27)
shannon <- as.data.frame(diversity(divmat))
colnames(shannon) <- "shannon"
rich <- specnumber(divmat01)
divdf <- shannon %>% 
  mutate(richness=rich,
         zone=zone,
         seas=seas)


```

Figure 2
```{r}

# Visualize: Specify the comparisons you want
my_comparisons <- list( c("Z1", "Z2"), c("Z2", "Z3"), c("Z1", "Z3") )
ggboxplot(divdf, x = "zone", y = "richness",
          color = "zone", palette = "jco")+ 
  stat_compare_means(comparisons = my_comparisons)+
  theme_bw() + scale_color_fivethirtyeight(name="Zone",labels=c("Forest","Disturbed","Cleared")) +# Add pairwise comparisons p-value
  stat_compare_means(label.y = 7.5,label.x = 0.7)+ xlab("Zone")+ ylab("Species richness")+
  scale_x_discrete(labels=c("Forest","Disturbed","Cleared"))+
  theme(legend.position = "none")# Add global p-value


```


How many times was each trap surveyed (number occasions excluding trap failure)=denominator of binomial
```{r}
table(ms_nsurveys <- apply(ms_sp_a[,,,1], c(1,2), function(x) sum(!is.na(x))))
```

Collapse 4D detection/nondetection data to 3D detection frequencies
Collapse to detection frequency per site and per trap=numerator of binomial
```{r}
ms_sp_asum <- apply(ms_sp_a, c(1,2,4), sum, na.rm = T)
```




# Model7: random effects at site per species, random slope of zone and season per species
occurrence species k at trap j of site s = intrercept[k] + re[k,s] x site[s] + rslopezone[k,j]x zone[k,j] + rslopeseas[k,j] x seas[k,j]
## Bundle and summarize data set
```{r}
season<-rep(0:1,times=9,each=3)

ZONE <- rep(1:3,times=18)

site <- rep(1:9,each=6)
```


```{r}
str( win.data <- list(ms_sp_asum = ms_sp_asum, M = nrow(ms_sp_asum), J = ms_nsurveys, zone=ZONE, season=season,site=site, nspec = dim(ms_sp_asum)[3],n.traps=50,nsite=9) )
setwd("C:\\Users\\mpruvot\\Desktop\\R Projects\\LUCms")

# Specify model in BUGS language
sink("model7.txt")
cat("
model {

# Priors for heterogeneity of species in community
for(k in 1:nspec){          # Loop over species
  lpsi[k] ~ dnorm(mu.lpsi,tau.lpsi)              #random species intercept
  for(s in 1:nsite){
    lpsi.site[k,s]~dnorm(0,tau.lpsi.site[k])
  }
  betalpsi1[k]~dnorm(mu.betalpsi1,tau.betalpsi1)
  betalpsi2[k]~dnorm(mu.betalpsi2,tau.betalpsi2)
  betalpsi.seas[k]~dnorm(mu.betalpsi.seas,tau.betalpsi.seas)
  betaint.zoneseas.lpsi1[k]~dnorm(mu.betaint.zoneseas.lpsi1,tau.betaint.zoneseas.lpsi1)
   betaint.zoneseas.lpsi2[k]~dnorm(mu.betaint.zoneseas.lpsi2,tau.betaint.zoneseas.lpsi2)
  lp[k] ~ dnorm(mu.lp,tau.lp)
}
#hyperpriors for full community

mu.lpsi ~ dnorm(0,0.1) # Community mean of occupancy (logit)
tau.lpsi <- pow(sd.lpsi, -2)
sd.lpsi ~ dunif(0,8) # Species heterogeneity in logit(psi)
for(k in 1:nspec){
tau.lpsi.site[k] <- pow(sd.lpsi.site[k], -2)
sd.lpsi.site[k]~dunif(0,8)
}
mu.betalpsi1~dnorm(0,0.01)
sd.betalpsi1~dunif(0,8)
tau.betalpsi1 <- pow(sd.betalpsi1, -2)
mu.betalpsi2~dnorm(0,0.01)
sd.betalpsi2~dunif(0,8)
tau.betalpsi2 <- pow(sd.betalpsi2, -2)
mu.betalpsi.seas~dnorm(0,0.01)
sd.betalpsi.seas~dunif(0,8)
tau.betalpsi.seas <- pow(sd.betalpsi.seas, -2)
mu.betaint.zoneseas.lpsi1~dnorm(0,0.01)
tau.betaint.zoneseas.lpsi1<- pow(sd.betaint.zoneseas.lpsi1, -2)
sd.betaint.zoneseas.lpsi1~dunif(0,8)
mu.betaint.zoneseas.lpsi2~dnorm(0,0.01)
tau.betaint.zoneseas.lpsi2<- pow(sd.betaint.zoneseas.lpsi2, -2)
sd.betaint.zoneseas.lpsi2~dunif(0,8)
mu.lp ~ dnorm(0,0.01) # Community mean of detection (logit)
tau.lp <- pow(sd.lp, -2)
sd.lp ~ dunif(0,8) # Species heterogeneity in logit(p)


# Ecological model for latent occurrence z (process model)
##model pecie k
for(k in 1:nspec){          # Loop over species

  ## Occurence in trap j of site i
  for (i in 1:M) {         # Loop over M sites/zone/season
     for(j in 1:n.traps){ logit(psi[i,j,k])<-lpsi[k]+lpsi.site[k,site[i]]+betalpsi1[k]*equals(zone[i],2)+betalpsi2[k]*equals(zone[i],3)+betalpsi.seas[k]*season[i] + betaint.zoneseas.lpsi1[k]*equals(zone[i],2)*season[i]+betaint.zoneseas.lpsi2[k]*equals(zone[i],3)*season[i]
      
      z[i,j,k] ~ dbern(psi[i,j,k])
        
  
           
        }
   }
}

#observation model

for(k in 1:nspec){          # Loop over species
logit(p[k])<-lp[k]
   for (i in 1:M) {         # Loop over sites
        for(j in 1:n.traps){ #loop over traps
            ms_sp_asum[i,j,k] ~ dbin(mup[i,j,k], J[i,j])
            mup[i,j,k] <- z[i,j,k] * p[k]
           
        }
   }
}


            

# Derived quantities
for(k in 1:nspec){          # Loop over species
for (i in 1:M) { 
temp[i,k] <- (sum(z[i,,k])>0)*1 #creates a temporary matrix of M*k with number of traps occupied
   }   # Add up number of occurring species at each site
}
for(k in 1:nspec){          # Loop over species
   Nocc.fs[k] <- sum(temp[,k]) # Add up number of occupied sites among the 27
}
for (i in 1:M) {            # Loop over sites
   Nsite[i] <- sum(temp[i,])   # Add up number of occurring species at each site
}

}
",fill = TRUE)
sink()

# Initial values
ms_zst <- apply(ms_sp_a, c(1,2,4), max, na.rm=T) # Observed occurrence as inits for z
ms_zst[is.na(ms_zst)] <- 1
inits <- function() list(z = ms_zst,  lpsi = rnorm(n= nspec),lpsi.site =matrix(rnorm(n=9*nspec),ncol=9),betalpsi1=rnorm(n=nspec),betalpsi2=rnorm(n=nspec),betalpsi.seas=rnorm(n=nspec),betaint.zoneseas.lpsi1=rnorm(n=nspec),betaint.zoneseas.lpsi2=rnorm(n=nspec), lp = rnorm(n= nspec))

# Parameters monitored
params <- c("mu.lpsi","sd.lpsi","mu.betalpsi1","sd.betalpsi1","mu.betalpsi2","sd.betalpsi2","lpsi","lpsi.site","betalpsi1","betalpsi2","mu.lp","sd.lp","psi","mu.betalpsi.seas","sd.betalpsi.seas","betalpsi.seas","betaint.zoneseas.lpsi1","betaint.zoneseas.lpsi2","sd.betaint.zoneseas.lpsi1","sd.betaint.zoneseas.lpsi2", "p", "Nsite", "Nocc.fs","lp","z")

params <- c("betalpsi1","betalpsi2","betalpsi.seas","betaint.zoneseas.lpsi1","betaint.zoneseas.lpsi2", "p", "Nsite", "Nocc.fs")

# MCMC settings
ni <- 300000   ;   nt <- 2   ;   nb <- 45000   ;   nc <- 3

# Call JAGS from R (ART 2.1 min)
out7 <- jags(win.data, inits, params, "model7.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE,n.cores = 8)

jagsUI::traceplot(out7)   

print(out7, dig = 3)



```



```{r}
rownames(out7$summary)
N.pm <- out7$summary[85:138, 1]       # Posterior means of Nsite
N.psd <- out7$summary[85:138, 2]      # ... posterior sd's of Nsite
N.cri <- out7$summary[85:138, c(3,7)] # ... CRL's of Nsite
  
  # Plot estimates as a function of zone (this does not propagate the uncertainty around N.pm estimates)

zone1 <- as.numeric(rep(1:3,times=18)==1)
zone2 <- as.numeric(rep(1:3,times=18)==2)
zone3 <- as.numeric(rep(1:3,times=18)==3)
season <- rep(c(0,1),each=9*3)


    pred.season <- c(0,0,0,1,1,1) 
  pred.zone <-  c(1,2,3,1,2,3)
  str(meta7.data <- list(zone2 = zone2, zone3=zone3, season=season, N = N.pm, psd = N.psd+0.1, pred.season=pred.season,pred.zone=pred.zone, n = length(N.pm),npred=length(pred.season))) ## the +0.1 was just to fix issues with psd values, but needs a more systemic solution
  
  setwd("C:\\Users\\mpruvot\\Desktop\\R Projects\\LUCms")

  # Define model in BUGS language
  sink("meta7.txt")
  cat("
model{

# Priors
for(v in 1:6){         # Priors for intercept and coefficients
   beta[v] ~ dnorm(0, 0.0001)
}
tau.site <- pow(sd.site, -2)
sd.site ~ dunif(0,10)

# Likelihood
for (i in 1:n){
   N[i] ~ dnorm(muN[i], tau.psd[i]) # Measurement error model for estimated N
   tau.psd[i] <- pow(psd[i], -2)    # 'Known' part of residual: meas. error
   muN[i] <- beta[1] + beta[2] * zone2[i] + beta[3] * zone3[i] + beta[4]*season[i] + beta[5] * season[i] * zone2[i] + beta[6] * season[i] * zone3[i] +eps.site[i] # add another source of uncertainty
   eps.site[i] ~ dnorm(0, tau.site) # this is the usual 'residual'
}

# Get predictions for plot

for(i in 1:npred){
   Npred[i] <- beta[1] + beta[2] * equals(pred.zone[i],2) + beta[3] * equals(pred.zone[i],3) + beta[4]*pred.season[i] + beta[5] * pred.season[i] * equals(pred.zone[i],2) + beta[6] * season[i] * equals(pred.zone[i],3)
}
} # end model
",fill=TRUE)
  sink()
  
  # Initial values, params monitored, and MCMC settings
  inits <- function() list(beta = rnorm(6))
  params <- c("beta", "sd.site", "Npred")
  ni <- 10000   ;   nt <- 10   ;   nb <- 1000   ;   nc <- 3
  
  # Call JAGS and summarize posterior
  outmeta7 <- jags(meta7.data, inits, params, "meta7.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb,parallel = TRUE,n.cores = 8)
  as.data.frame(outmeta7$summary) %>% rownames_to_column() %>% flextable() %>% colformat_double(digits = 3) %>%  save_as_docx(path=file.choose())
  str(outmeta7$summary)
```




# Figure 4
```{r}

outgraph <-  as.data.frame(outmeta7$summary[8:13,])
outgraph$zone <- c("Z1","Z2","Z3","Z1","Z2","Z3")
outgraph$season <- c("d","d","d","r","r","r")
colnames(outgraph)[c(3,7)] <- c("low","high")


p.pred.df <- data.frame(Zone = factor(rep(c("zone1","zone2","zone3"),2)), #obs range
                        Season = factor(rep(c("d","r"),each=3)))
outgraph <- cbind(outgraph,p.pred.df)




ggplot(outgraph, aes(x=season, y=mean, color = zone, group=zone)) +
  geom_errorbar(aes(ymin=low,ymax=high), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=5, position = position_dodge(width=0.5)) +
  theme_bw() + scale_color_fivethirtyeight(name="Zone",labels=c("Forest","Disturbed","Cleared")) +
  xlab("Season") + ylab("Predicted Species Richness")+                
  scale_x_discrete(labels=c("Dry", "Rainy"))



```


