---
title: "oSCR analysis"
author: "Mathieu Pruvot"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r librairies}
library(tidyverse)
library(oSCR)
library(gt)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(flextable)
```

# Upload original data
## trap data

Load three versions of the trap layout:
- traplayour_tot2 only contains basic trap layout info
- traplayout_tot3 contains basic+veg transect
- traplayout_tot4 contains basic+veg transect+veg assessment

```{r load trap data}
load("traplayout_tot2.Rdata")
load("traplayout_tot3.Rdata")
load("traplayout_tot4.Rdata")


```




## Upload capture data

```{r load capt data}
load("~/Work/WCS/LACANET/Land-use study/Analysis/Analysis sept 2019/captanim22.Rdata")
head(captanim22)
colnames(captanim22)
captanim22$Sex <- factor(captanim22$Sex)

```
remove the -occasion for mortality events, but save a version in OccasionD
```{r}
captanim22$OccasionD <-captanim22$Occasion
captanim22$Occasion <- abs(captanim22$Occasion)
```



Re-order variables of encounter data file 
```{r}
captanim<-captanim22[,c(1,2,3,4,12,5:11,13:42)]
captanim$Sex <- (captanim$Sex=="Female")*1

```


Clean up traplayout file
```{r}
head(traplayout_tot4)
traplayout_tot4_oscr <- traplayout_tot4 %>% separate(usage,c("1","2","3","4","5","6","7","8"),sep=1:7,convert = T)

traplayout_tot4_oscr$Site.x <- str_remove(traplayout_tot4_oscr$Site.x,"[/]")
traplayout_tot4_oscr$sep <- "/"
colnames(traplayout_tot4_oscr)
traplayout_tot4_oscr <- traplayout_tot4_oscr[,c(1:11,99,12:98)]
head(traplayout_tot4_oscr)
traplayout_tot4_oscr$session<-substr(traplayout_tot4_oscr$TrapID.x,1,5)
colnames(traplayout_tot4_oscr)[4:11] <- paste0("O.",colnames(traplayout_tot4_oscr)[4:11])

```
Decriptive statistics on captures
```{r}

captnigth <- sum(traplayout_tot4_oscr[,4:11])
totacapt <- nrow(captanim)
uniqueid <- length(unique(captanim$Id))
recapevents <- totacapt-uniqueid
totacapt
uniqueid
recapevents
```
Total of 20214 trap-night after excluding trap mal-function.

```{r}
captanim %>% 
  group_by(Id) %>% 
  summarize(n=n()) %>% 
  summarize(recap=sum(n>1), norecap=sum(n==1), tot=n(), recaprate=recap/tot)
```
capture rate
```{r}
captrate <- totacapt/captnigth 
newcaptrate <- uniqueid/captnigth
recaptrate <- recapevents/captnigth
captrate
newcaptrate
recaptrate
```

```{r}
uniqueanimaldata <- captanim %>% 
  group_by(Id) %>% 
  slice(1)

levels(uniqueanimaldata$species) <- list('Bandicota savilei'="Bandicotasavilei",'Berylmys berdmorei'="Berylmysberdmorei", 'Berylmys bowersi'="Berylmysbowersi",'Chiropodomys gliroides'="Chiropodomysgliroides",'Leopoldamys sabanus'="Leopoldamyssabanus",'Maxomys surifer'="Maxomyssurifer", 'Mus caroli'="Muscaroli",'Mus cervicolor'="Muscervicolor",'Niviventer fulvescens'="Niviventerfulvescens",'Rattus sp. R3'="RattusspR3",'Rattus andamanensis'="Rattusandamanensis",'Rattus exulans'="Rattusexulans",'Tupaia belangeri'="Tupaiabelangeri",'Vandeleuria oleracea'="Vandeleuriaoleracea")

uniqueanimaldata$species <- as.character(uniqueanimaldata$species)

uniqueanimaldata %>% 
  group_by(species) %>% 
  summarize(tot=n(), prop=tot/965) %>% 
  arrange(-prop)

tbl1 <- sort(table(uniqueanimaldata$species),decreasing = T)
tbl2 <- prop.table(tbl1)
cbind(Freq=tbl1,Proportion=tbl2,Cumul=cumsum(tbl2))
```

Table 2
```{r}


df1 <- uniqueanimaldata %>% 
  select(Season.x,Zone.x,species) 
df1 <- df1[,2:4]

tbr <- df1%>% 
  filter(Season.x=="r") %>% 
 tbl_summary( by=Zone.x, include=c(species),sort=list(everything() ~ "frequency"),statistic = list(species ~ "{n}") ) %>% 
  add_overall(
  col_label = "**Total Rainy**",
  statistic = ~ "{n} ({p}%)",
  last = T
)%>% 
   modify_header(label ~ "**Species**",
                 stat_1 ~ "**Forest**, N = 68",
                 stat_2 ~ "**Disturbed**, N = 136",
                  stat_3 = "**Cleared**, N = 387") %>% 
  remove_row_type(
  variables = species,
  type = c("header")
) %>% 
  italicize_levels()


tbd <- df1%>% 
  filter(Season.x=="d") %>% 
 tbl_summary( by=Zone.x, include=c(species),sort=list(everything() ~ "frequency"),statistic = list(species ~ "{n}"))%>%   add_overall(
  col_label = "**Total Dry**",
  statistic = ~ "{n} ({p}%)",
  last = T,
) %>% 
   modify_header(label ~ "**Species**",
                 stat_1 ~ "**Forest**, N = 91",
                 stat_2 ~ "**Disturbed**, N = 125",
                  stat_3 = "**Cleared**, N = 158") %>% 
  remove_row_type(
  variables = species,
  type = c("header")
)%>% 
  italicize_levels()

tbtot <- df1%>% 
 tbl_summary(  include=c(species),sort=list(everything() ~ "frequency")) %>% 
  remove_row_type(
  variables = species,
  type = c("header")
)%>% 
   modify_header(label ~ "**Species**")%>% 
  italicize_levels()

theme_gtsummary_compact()
tbl_merge(
  list(tbtot,tbr, tbd),
  tab_spanner = c("**Total**","**Rainy Season**", "**Dry Season**")
) %>%  as_flex_table() %>% 
  save_as_docx(path="C:\\Users\\mpruvot\\Desktop\\R Projects\\LUCms\\table1.docx")

```

# Data preparation for oSCR analysis on all-rodent data


Prepare tdf
```{r}
colnames(traplayout_tot4_oscr)[13:100] <- gsub("\\.","",colnames(traplayout_tot4_oscr)[13:100])
covnames <- colnames(traplayout_tot4_oscr)[c(14,16:61,70,71,74,76,77,79:81,83,85,86,87,92,97,100)]
tdflist<-split(traplayout_tot4_oscr,traplayout_tot4_oscr$session)
tdflist <- tdflist[names(tdflist) != "S7dZ1"]
```


- remove site S7dZ1 from previously defined objectives in rodent.data to add variables to trapCov
```{r}
oscrsite <- paste0("site",site[-37])
oscrseason <- rep(c("d","r"),times=9,each=3)[-37]
oscrzone <- paste0("zone",zone[-37])

```
Add site, season and zone to tdf so it can be used as trap covariate for modeling Po/Go
```{r}

for(i in 1:53){

  tdflist[[i]]$site <- factor(oscrsite[i],levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9"))
    tdflist[[i]]$season <- factor(oscrseason[i],levels=c("d","r"))
      tdflist[[i]]$zone <- factor(oscrzone[i],levels=c("zone1","zone2","zone3"))
}

covnames <- c("zone",covnames)
```

Add day and day2
```{r}
for(i in 1:53){

    tdflist[[i]]$day.1 <-0
      tdflist[[i]]$day.2 <-1
        tdflist[[i]]$day.3 <-2
          tdflist[[i]]$day.4 <-3
            tdflist[[i]]$day.5 <-4
              tdflist[[i]]$day.6 <-5
                tdflist[[i]]$day.7 <-6
                  tdflist[[i]]$day.8 <-7

    tdflist[[i]]$day2.1 <-0
      tdflist[[i]]$day2.2 <-1
        tdflist[[i]]$day2.3 <-4
          tdflist[[i]]$day2.4 <-9
            tdflist[[i]]$day2.5 <-16
              tdflist[[i]]$day2.6 <-25
                tdflist[[i]]$day2.7 <-36
                  tdflist[[i]]$day2.8 <-49


}

covnames <- c("day","day2", covnames)
```


Create the scrFrame
```{r}
msdata <- data2oscr(edf = captanim,
tdf = tdflist,
sess.col = which(colnames(captanim) %in% "session"),
id.col = which(colnames(captanim) %in% "Id"),
occ.col = which(colnames(captanim) %in% "Occasion"),
trap.col = which(colnames(captanim) %in% "TrapID"),
sex.col= which(colnames(captanim) %in% "Sex"),
sex.nacode = "NA",
K = rep(8,53),
ntraps = rep(50,53),
trapcov.names = covnames)


rod.sf <- msdata$scrFrame


rod.sf
plot(rod.sf)

```



prepare an addition to indcovs: add 'removed' variable 

first create the remove variable in the captanim data
```{r}
captanim$removed <- (captanim$OccasionD<0)*abs(captanim$OccasionD)
captanim$removed[captanim$removed==0]<-8
```

then convert captanim$removed to the indCovs format (code from Chris Sutherland on oSCR google group)
```{r}
for(i in 1:53){
tmp <- captanim[captanim$session == names(tdflist)[i],]
unique(tmp$Id)
tmp$Id <- factor(tmp$Id)
rmv <- tapply(tmp$removed,tmp$Id,mean)
rod.sf$indCovs[[i]]$removed <- rmv
}

```


```{r}
rod.ss <- make.ssDF(scrFrame = rod.sf, 
buffer = 3*23, # ~3*1/2*MMDM
res = 5)
plot.ssDF(rod.ss)
plot(rod.ss,rod.sf)
```
Follow instruction from CHris Sutherland on adding state space covariates to model covariate effet on 
- Density:
```{r}
for(i in 1:53){

  rod.ss[[i]]$site <- factor(oscrsite[i],levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9"))
    rod.ss[[i]]$season <- factor(oscrseason[i],levels=c("d","r"))
      rod.ss[[i]]$zone <- factor(oscrzone[i],levels=c("zone1","zone2","zone3"))

}

```

- Sigma
```{r}
rod.sf$sigCovs$site <- factor(oscrsite,levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9")) 
rod.sf$sigCovs$season <- factor(oscrseason,levels=c("d","r"))
rod.sf$sigCovs$zone <- factor(oscrzone,levels=c("zone1","zone2","zone3"))


```


Create a version of rod.sf wthout the sex variable
```{r}
rod.sf2 <- rod.sf
rod.sf2$indCovs <- lapply(rod.sf2$indCovs, function(x) x[!(names(x) %in% c("sex"))])
```

Run the all-species models without sex and with multi-core (this was run a remote server)
```{r}
mods <- make.mods(density= c(~1,~zone,~season,~zone+season,~zone*season),
                  detection = c(~1,~zone,~season,~zone+season,~zone*season),
                  sigma = c(~1,~zone),
)

library(foreach)

library(doParallel)



cl <- makeCluster(25)    #make the cluster

registerDoParallel(cl)  #register the cluster



out2 <- foreach(i=1:50,.packages = "oSCR",.errorhandling="pass") %dopar% {
  
  
  
  m <- list(mods[i,1][[1]], # ith model
            
            mods[i,2][[1]],
            
            mods[i,3][[1]],
            
            mods[i,4][[1]]) 
  
  out <- oSCR.fit(m, rod.sf2,rod.ss)
  
  return(out)
  
}

stopCluster(cl)
convmod <- summary(out2)[,1]=="17"
out2oscr <- fitList.oSCR(out2[convmod], rename=T)

namod <- numeric()

for (i in 1:length(out2oscr)){
  if(anyNA(out2oscr[[i]]$outStats$std.er)){namod <- c(namod,i)}
}
out2oscr <- fitList.oSCR(out2oscr[-namod], rename=T)
modsel <- modSel.oSCR(out2oscr)


```

Suppl table
```{r}
library(flextable)
modsel$aic.tab %>% flextable() %>% colformat_double(digits = 3) 
```

pick the best all-rodent model with the lowest AIC
```{r}
topmod <- out2oscr$`D(~zone + season + zone:season) p(~zone + season) sig(~1) asu(~1)`
```

make a dataframe of values for predictions of all-rodent density
```{r}
p.pred.df <- data.frame(zone = factor(rep(c("zone1","zone2","zone3"),2)), #obs range
                        season = factor(rep(c("d","r"),each=3)))
library(oSCR)
#now predict
p.preds <- get.real(model = topmod, type = "det", newdata = p.pred.df)
d.preds <- get.real(model = topmod, type = "dens", newdata = p.pred.df,d.factor = 1/25)
s.preds <- get.real(model = topmod, type = "sig", newdata = p.pred.df)
```

Figure 3a
```{r}
ggplot(d.preds, aes(x=season, y=estimate, color = zone, group=zone)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=5, position = position_dodge(width=0.5)) +
  theme_bw() + scale_color_fivethirtyeight(name="Zones",labels=c("Forest","Disturbed","Cleared")) +
  xlab("Season") + ylab(expression("Density in animal /"~m^2))+
  scale_x_discrete(labels=c("Dry", "Rainy"))

```


# prepare M surifer data oSCR analysis 


```{r}
captanim.MS <- captanim[captanim$species=="Maxomyssurifer",]
MS.tdfname <- unique(paste0(captanim.MS$Site.x,captanim.MS$Season.x,captanim.MS$Zone.x))
```


```{r}
tdflist.MS <- tdflist[names(tdflist) %in% MS.tdfname]


oscrsite.MS <- c(rep("site1",4),rep("site2",3),rep("site3",3),rep("site4",4),rep("site5",4),rep("site6",3),rep("site7",2),rep("site8",4),rep("site9",4))
oscrseason.MS <- c("d","d","r","r","d","d","r","d","r","r","d","d","r","r","d","d","r","r","d","d","r","r","r","d","d","r","r","d","d","r","r")
oscrzone.MS <- str_replace(str_sub(names(tdflist.MS),4,5),"Z","zone")
```


Create the scrFrame
```{r}
msdata.MS <- data2oscr(edf = captanim.MS,
tdf = tdflist.MS,
sess.col = which(colnames(captanim.MS) %in% "session"),
id.col = which(colnames(captanim.MS) %in% "Id"),
occ.col = which(colnames(captanim.MS) %in% "Occasion"),
trap.col = which(colnames(captanim.MS) %in% "TrapID"),
sex.col= which(colnames(captanim.MS) %in% "Sex"),
sex.nacode = "NA",
K = rep(8,31),
ntraps = rep(50,31),
trapcov.names = covnames)


rod.MS.sf <- msdata.MS$scrFrame


rod.MS.sf
plot(rod.MS.sf)

```






```{r}
rod.MS.ss <- make.ssDF(scrFrame = rod.MS.sf, # the RBS MS scrFrame
buffer = 3*23, # ~3*1/2*MMDM
res = 5)
plot.ssDF(rod.MS.ss)
plot(rod.MS.ss,rod.MS.sf)
```
Follow instruction from CHris Sutherland on adding state space covariates to model covariate effet on 
- Density:
```{r}
for(i in 1:31){

  rod.MS.ss[[i]]$site <- factor(oscrsite.MS[i],levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9"))
    rod.MS.ss[[i]]$season <- factor(oscrseason.MS[i],levels=c("d","r"))
      rod.MS.ss[[i]]$zone <- factor(oscrzone.MS[i],levels=c("zone1","zone2","zone3"))

}

```

- Sigma
```{r}
rod.MS.sf$sigCovs$site <- factor(oscrsite.MS,levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9")) 
rod.MS.sf$sigCovs$season <- factor(oscrseason.MS,levels=c("d","r"))
rod.MS.sf$sigCovs$zone <- factor(oscrzone.MS,levels=c("zone1","zone2","zone3"))


```


# Prepare Mus cervicolor data for oSCR analysis 

```{r}
captanim.MC <- captanim[captanim$species=="Muscervicolor",]
MC.tdfname <- unique(paste0(captanim.MC$Site.x,captanim.MC$Season.x,captanim.MC$Zone.x))
```


```{r}
tdflist.MC <- tdflist[names(tdflist) %in% MC.tdfname]


oscrsite.MC <- str_replace(str_sub(names(tdflist.MC),1,2),"S","site")
oscrseason.MC <- str_sub(names(tdflist.MC),3,3)
oscrzone.MC <- str_replace(str_sub(names(tdflist.MC),4,5),"Z","zone")
```


Create the scrFrame
```{r}
msdata.MC <- data2oscr(edf = captanim.MC,
tdf = tdflist.MC,
sess.col = which(colnames(captanim.MC) %in% "session"),
id.col = which(colnames(captanim.MC) %in% "Id"),
occ.col = which(colnames(captanim.MC) %in% "Occasion"),
trap.col = which(colnames(captanim.MC) %in% "TrapID"),
sex.col= which(colnames(captanim.MC) %in% "Sex"),
sex.nacode = "NA",
K = rep(8,24),
ntraps = rep(50,24),
trapcov.names = covnames)


rod.MC.sf <- msdata.MC$scrFrame


rod.MC.sf
plot(rod.MC.sf)

```






```{r}
rod.MC.ss <- make.ssDF(scrFrame = rod.MC.sf, # the RBS MS scrFrame
buffer = 3*15, # ~3*1/2*MMDM
res = 5)
plot.ssDF(rod.MC.ss)
plot(rod.MC.ss,rod.MC.sf)
```
Follow instruction from CHris sutherland on adding state space covariates to model covariate effet on 
- Density:
```{r}
for(i in 1:24){

  rod.MC.ss[[i]]$site <- factor(oscrsite.MC[i],levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9"))
    rod.MC.ss[[i]]$season <- factor(oscrseason.MC[i],levels=c("d","r"))
      rod.MC.ss[[i]]$zone <- factor(oscrzone.MC[i],levels=c("zone1","zone2","zone3"))

}

```

- Sigma
```{r}
rod.MC.sf$sigCovs$site <- factor(oscrsite.MC,levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9")) 
rod.MC.sf$sigCovs$season <- factor(oscrseason.MC,levels=c("d","r"))
rod.MC.sf$sigCovs$zone <- factor(oscrzone.MC,levels=c("zone1","zone2","zone3"))


```

# Prepare Rattus sp R3 data for oSCR analysis  

```{r}
captanim.RR <- captanim[captanim$species=="RattusspR3",]
RR.tdfname <- unique(paste0(captanim.RR$Site.x,captanim.RR$Season.x,captanim.RR$Zone.x))
```


```{r}
tdflist.RR <- tdflist[names(tdflist) %in% RR.tdfname]


oscrsite.RR <- str_replace(str_sub(names(tdflist.RR),1,2),"S","site")
oscrseason.RR <- str_sub(names(tdflist.RR),3,3)
oscrzone.RR <- str_replace(str_sub(names(tdflist.RR),4,5),"Z","zone")


```


Create the scrFrame
```{r}
msdata.RR <- data2oscr(edf = captanim.RR,
tdf = tdflist.RR,
sess.col = which(colnames(captanim.RR) %in% "session"),
id.col = which(colnames(captanim.RR) %in% "Id"),
occ.col = which(colnames(captanim.RR) %in% "Occasion"),
trap.col = which(colnames(captanim.RR) %in% "TrapID"),
sex.col= which(colnames(captanim.RR) %in% "Sex"),
sex.nacode = "NA",
K = rep(8,37),
ntraps = rep(50,37),
trapcov.names = covnames)


rod.RR.sf <- msdata.RR$scrFrame


rod.RR.sf
plot(rod.RR.sf)

```






```{r}
rod.RR.ss <- make.ssDF(scrFrame = rod.RR.sf, # the RBS MS scrFrame
buffer = 3*23, # ~3*1/2*MMDM
res = 5)
plot.ssDF(rod.RR.ss)
plot(rod.RR.ss,rod.RR.sf)
```
Follow instruction from CHris sutherland on adding state space covariates to model covariate effet on 
- Density:
```{r}
for(i in 1:37){

  rod.RR.ss[[i]]$site <- factor(oscrsite.RR[i],levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9"))
    rod.RR.ss[[i]]$season <- factor(oscrseason.RR[i],levels=c("d","r"))
      rod.RR.ss[[i]]$zone <- factor(oscrzone.RR[i],levels=c("zone1","zone2","zone3"))

}

```

- Sigma
```{r}
rod.RR.sf$sigCovs$site <- factor(oscrsite.RR,levels=c("site1","site2","site3","site4","site5","site6","site7","site8","site9")) 
rod.RR.sf$sigCovs$season <- factor(oscrseason.RR,levels=c("d","r"))
rod.RR.sf$sigCovs$zone <- factor(oscrzone.RR,levels=c("zone1","zone2","zone3"))


```
Create a version of rod.sf wthout the sex variable
```{r}
rod.MS.sf2 <- rod.MS.sf
rod.MS.sf2$indCovs <- lapply(rod.MS.sf2$indCovs, function(x) x[!(names(x) %in% c("sex"))])

rod.MC.sf2 <- rod.MC.sf
rod.MC.sf2$indCovs <- lapply(rod.MC.sf2$indCovs, function(x) x[!(names(x) %in% c("sex"))])

rod.RR.sf2 <- rod.RR.sf
rod.RR.sf2$indCovs <- lapply(rod.RR.sf2$indCovs, function(x) x[!(names(x) %in% c("sex"))])
```

For each predominant species, run models without sex and with multi-core
```{r}
mods <- make.mods(density= c(~1,~zone,~season,~zone+season,~zone*season),
                  detection = c(~1,~zone,~season,~zone+season,~zone*season),
                  sigma = c(~1,~zone,~season),
)

library(foreach)

library(doParallel)



cl <- makeCluster(8)    #make the cluster

registerDoParallel(cl)  #register the cluster



out2.MS <- foreach(i=1:75,.packages = "oSCR",.errorhandling="pass") %dopar% {
  
  
  
  m <- list(mods[i,1][[1]], # ith model
            
            mods[i,2][[1]],
            
            mods[i,3][[1]],
            
            mods[i,4][[1]]) 
  
  
  out <- oSCR.fit(m, rod.MS.sf2,rod.MS.ss)
  
  return(out)
  
}

out2.MC <- foreach(i=1:75,.packages = "oSCR",.errorhandling="pass") %dopar% {
  
  
  
  m <- list(mods[i,1][[1]], # ith model
            
            mods[i,2][[1]],
            
            mods[i,3][[1]],
            
            mods[i,4][[1]]) 
  
  
  out <- oSCR.fit(m, rod.MC.sf2,rod.MC.ss)
  
  return(out)
  
}

out2.RR <- foreach(i=1:75,.packages = "oSCR",.errorhandling="pass") %dopar% {
  
  m <- list(mods[i,1][[1]], # ith model
            
            mods[i,2][[1]],
            
            mods[i,3][[1]],
            
            mods[i,4][[1]]) 
  
  out <- oSCR.fit(m, rod.RR.sf2,rod.RR.ss)
  
  return(out)
  
}

stopCluster(cl)


```

# Species-specific prediction plots and tables

Selection best model for Maxomys surifer
```{r}
library(oSCR)
modsel.MS <- modSel.oSCR(out2oscr.MS)
modsel.MS$aic.tab %>% flextable() %>% colformat_double(digits = 3) %>%  save_as_docx(path=file.choose())

#pick the model with the lowest AIC
topmod.MS <- out2oscr.MS$`D(~zone) p(~season) sig(~season) asu(~1)`
```

make a dataframe of values for predictions for Maxomys surifer
```{r}

p.pred.df <- data.frame(zone = factor(rep(c("zone1","zone2"),2),levels = c("zone1","zone2")), #obs range
                        season = factor(rep(c("d","r"),each=2)))
#now predict
p.preds.MS <- get.real(model = topmod.MS, type = "det", newdata = p.pred.df, d.factor = 1/25)
d.preds.MS <- get.real(model = topmod.MS, type = "dens", newdata = p.pred.df, d.factor = 1/25)
s.preds.MS <- get.real(model = topmod.MS, type = "sig", newdata = p.pred.df, d.factor = 1/25)

head(p.preds.MS)
levels(d.preds.MS$zone) <- c("zone1","zone2","zone3")
d.preds.MS <- d.preds.MS %>% add_row(zone = "zone3", season = "d",estimate=0) %>% add_row(zone = "zone3", season = "r",estimate=0,)
```

Figure 3b
```{r}
ggplot(d.preds.MS, aes(x=season, y=estimate, color = zone, group=zone)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=5, position = position_dodge(width=0.5)) +
  theme_bw() + scale_color_fivethirtyeight(name="Zone",labels=c("Forest","Disturbed","Cleared")) +
  xlab("Season") + ylab(expression("Density Maxomys surifer in animal / "~m^2))+
  scale_x_discrete(labels=c("Dry", "Rainy"))





```
## Select best model Mus cervicolor
```{r}
library(oSCR)
namod <- numeric()

for (i in 1:length(out2oscr.MC)){
  if(anyNA(out2oscr.MC[[i]]$outStats$std.er)){namod <- c(namod,i)}
}
out2oscr.MC <- fitList.oSCR(out2oscr.MC[-namod], rename=T)

modsel.MC <- modSel.oSCR(out2oscr.MC)
```

Suppl table
```{r}
modsel.MC$aic.tab %>% flextable() %>% colformat_double(digits = 3) 


```

pick the model with the lowest AIC
```{r}
topmod.MC <- out2oscr.MC$`D(~zone + season + zone:season) p(~season) sig(~season) asu(~1)`
```

make a dataframe of values for DETECTION predictions
```{r}

p.pred.df <- data.frame(zone = factor(rep(c("zone2","zone3"),2),levels = c("zone2","zone3")), #obs range
                        season = factor(rep(c("d","r"),each=2)))
```

now predict density of M cervicolor across zones and seasons
```{r}

p.preds.MC <- get.real(model = topmod.MC, type = "det", newdata = p.pred.df, d.factor = 1/25)
d.preds.MC <- get.real(model = topmod.MC, type = "dens", newdata = p.pred.df, d.factor = 1/25)
s.preds.MC <- get.real(model = topmod.MC, type = "sig", newdata = p.pred.df, d.factor = 1/25)

head(p.preds.MC)
levels(d.preds.MC$zone) <- c("zone2","zone3","zone1")
d.preds.MC <- d.preds.MC %>% add_row(zone = "zone1", season = "d",estimate=0) %>% add_row(zone = "zone1", season = "r",estimate=0,)

```

Figure 3d
```{r}
ggplot(d.preds.MC, aes(x=season, y=estimate, color = zone, group=zone)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=5, position = position_dodge(width=0.5)) +
  theme_bw() + scale_color_fivethirtyeight(name="Zone",labels=c("Forest","Disturbed","Cleared")) +
  xlab("Season") + ylab(expression("Density Mus cervicolor in animal / "~m^2))+
  scale_x_discrete(labels=c("Dry", "Rainy"))

```
Predctions and plots for Rattus sp R3
```{r}
modsel.RR <- modSel.oSCR(out2oscr.RR)
```

Suppl doc table
```{r}
modsel.RR$aic.tab %>% flextable() %>% colformat_double(digits = 3)
```

pick top model with  lowest AIC
```{r}
topmod.RR <- out2oscr.RR$`D(~zone) p(~zone) sig(~season) asu(~1)`
```

make a dataframe of values for DETECTION predictions
```{r}

p.pred.df <- data.frame(zone = factor(rep(c("zone1","zone2","zone3"),2)), #obs range
                        season = factor(rep(c("d","r"),each=3)))
library(oSCR)
#now predict
p.preds.RR <- get.real(model = topmod.RR, type = "det", newdata = p.pred.df)
d.preds.RR <- get.real(model = topmod.RR, type = "dens", newdata = p.pred.df,d.factor = 1/25)
s.preds.RR <- get.real(model = topmod.RR, type = "sig", newdata = p.pred.df)


```

Figure 3c
```{r}
ggplot(d.preds.RR, aes(x=season, y=estimate, color = zone, group=zone)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=5, position = position_dodge(width=0.5)) +
  theme_bw() + scale_color_fivethirtyeight(name="Zone",labels=c("Forest","Disturbed","Cleared")) +
  xlab("Season") + ylab(expression("Density Rattus sp. R3 in animal / "~m^2))+
  scale_x_discrete(labels=c("Dry", "Rainy"))


```
